---
title: "Cell  Migration timeplapse quantification"
output: html_notebook
author: "Federico Saltarin"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
---
As this is an R-Notebook you can run the single chunks or you can also do run all after setting your parameters and get immediately your final results. After running you will also get a html format that shows you the code along with the final results (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).
<br> 
<br>
<br>
First we define all the needed libraries
```{r}
library(dplyr)
library(ggplot2)
library(data.table)
library(tcltk)
```
<br> 
<br>
Then define some variables: Time_Res to define our time resolution and Pixel_Size to define the pixel size in um
```{r}
Time_Res = 15
Pixel_Size = 1.24
```
<br> 
<br>
Here we ask the user which CSV file to open (with the cell migration data)
```{r}
Data <- read.csv(tk_choose.files(caption = "Select results file: "))
```
<br> 
<br>
Here we start the analysis:
* Unique_ID will identify a specific cell of 1 specific Condition 
* X_Shift and Y_Shift calculate the shift in X and Y, respectively, for each consecutive timepoint 
* Dist is the distance traveled per timepoint 
* Instant_Veloc is the istantaneous velocity (per each timepoint) 

```{r}
Data <- Data %>%
        mutate(Unique_ID = paste0(Condition, ID)) %>%
        mutate(Time = Slice*Time_Res/60) %>%
        group_by(Unique_ID) %>%
        mutate(X_Shift = c(0,diff(X))^2,
               Y_Shift = c(0,diff(Y))^2,
               Dist = sqrt(X_Shift + Y_Shift),
               Instant_Veloc = Dist/unique(Time[2]),
               X_Shift = X - X[1],
               Y_Shift = Y - Y[1])

```
<br>
<br>
* Tot_Dist is the total migration distance
* Mean_Speed is the average migration speed
* Persistence is the "directionality" of the cell -> 1 = totally linear 0 = very unlinear
```{r}
Data_agg <- Data %>%
            group_by(Unique_ID)%>%
            summarise(Condition = unique(Condition), 
                      Tot_Dist = sum(Dist, na.rm = T),
                      Mean_Speed = mean(Instant_Veloc, na.rm = T),
                      X_Displ = (first(X)-last(X))^2,
                      Y_Displ = (first(Y)-last(Y))^2,
                      Displ = sqrt(X_Displ + Y_Displ),
                      Persistence = (Displ/Tot_Dist))
```
<br>
<br>
Data_Avg_Speed will be used to compute the average speed, grouped per condition
x_limit and y_limit are used to define the limits for the full tracks plot

```{r}
Data_Avg_Speed <- Data %>%
  group_by(Time, Condition) %>% 
  summarise(Avg_Speed = mean(Instant_Veloc))

x_limit = ifelse(abs(min(Data$X_Shift)) > abs(max(Data$X_Shift)), abs(min(Data$X_Shift)), abs(max(Data$X_Shift)))
y_limit = ifelse(abs(min(Data$Y_Shift)) > abs(max(Data$Y_Shift)), abs(min(Data$Y_Shift)), abs(max(Data$Y_Shift)))
```
<br> 
<br>
Then we create all the plots
- Total distance and mean speed will show basically the same information but with different units
- Persistence will show the directionality of cell migration
- Instantaneous velocity shows how migration velocity changes over time
- Tracks_Lines_Centered shows the full track of all single cells, centered to the origin. This plot is useful to understand if the   migration is random or directed to a specific direction (persistence helps too). 
```{r}

Tot_Dist_Boxplot <- ggplot(Data_agg, aes(x = Condition, y = Tot_Dist,
                     fill = Condition)) +
                      geom_boxplot(notch = T, alpha=0.7) +
                      geom_jitter(height = 0, width = 0.05, alpha=0.7, size = 3) +
                      theme_bw()+
                      theme(axis.text.x = element_blank(),
                      axis.title.x = element_blank(),
                      axis.ticks.x = element_blank())+
                      ylab('Total Distance (um)')


Mean_Speed_Boxplot <- ggplot(Data_agg, aes(x = Condition, y = Mean_Speed, fill = Condition)) +
                      geom_boxplot(notch = T, alpha=0.7) +
                      geom_jitter(height = 0, width = 0.05, alpha=0.7, size = 3) +
                      theme_bw()+
                      theme(axis.text.x = element_blank(),
                      axis.title.x = element_blank(),
                      axis.ticks.x = element_blank())+
                      ylab('Mean Speed (um/min)')
                    

Persistence_Boxplot <- ggplot(Data_agg, aes(x = Condition, y = Persistence,
                                         fill = Condition)) +
                        geom_boxplot(notch = F, alpha=0.7) +
                        geom_jitter(height = 0, width = 0.05, alpha=0.7, size = 3) +
                        theme_bw()+
                        theme(axis.text.x = element_blank(),
                        axis.title.x = element_blank(),
                        axis.ticks.x = element_blank())+
                        ylab('Persistence')
                   

Inst_Veloc_Line <- ggplot(Data, aes(x = Time, y = Instant_Veloc))+ 
                    geom_line(aes(group = Unique_ID))+
                    geom_line(data=Data_Avg_Speed, aes(x = Time, y = Avg_Speed), color = 'red', size = 1.5)+
                    facet_wrap("Condition")+
                    theme_bw()+
                    xlab('Time (min)')+
                    ylab('Instantaneous Velocity (um/min)')
                    

Tracks_Lines_Centered <- ggplot(Data, aes(x = X_Shift, y = Y_Shift))+ 
                          geom_line(aes(group = Unique_ID), size = 0.3)+
                          geom_point(x=0,y=0, col = "red", size = 2)+
                          facet_wrap("Condition")+
                          xlim(-x_limit, x_limit)+
                          ylim(-y_limit, y_limit)+
                          theme_bw()+
                          xlab('X')+
                          ylab('Y')
```
<br>
<br>
Here we display all the plots
```{r}
Tot_Dist_Boxplot
Mean_Speed_Boxplot
Persistence_Boxplot
Inst_Veloc_Line
Tracks_Lines_Centered

```
<br>
<br>
And then, if we want (just uncomment the code), we save all the plots in a new folder in PDF format
```{r}
# dir.create("../Plots")
# setwd("../Plots")
# 
# 
# Plots_List = list(Tot_Dist_Boxplot,
#                   Mean_Speed_Boxplot,
#                   Persistence_Boxplot,
#                   Inst_Veloc_Line,
#                   Tracks_Lines_Centered)
# 
# names(Plots_List) <- c("Tot_Dist_Boxplot",
#                        "Mean_Speed_Boxplot",
#                        "Persistence_Boxplot",
#                        "Inst_Veloc_Line",
#                        "Tracks_Lines_Centered")
# lapply(names(Plots_List),
#        function(x)
#          ggsave(filename=paste(x,".pdf",sep=""), height = 7, width = 15, plot=Plots_List[[x]]))
```










