---
title: "Cell  Migration timeplapse quantification"
output: html_notebook
author: "Federico Saltarin"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
---
As this is an R-Notebook you can run the single chunks or you can also do run all after setting your parameters and get immediately your final results. After running you will also get a html format that shows you the code along with the final results (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).
<br> 
<br>
<br>
First we define all the needed libraries
```{r}
library(dplyr)
library(ggplot2)
library(data.table)
library(tcltk)
library(gganimate)
```
<br> 
<br>
Then define some variables: Time_Res to define our time resolution and Pixel_Size to define the pixel size in um
```{r}
Time_Res = 2
Pixel_Size = 1
```
<br> 
<br>
Here we ask the user which CSV file to open (with the cell migration data)
```{r}
Data <- read.csv(tk_choose.files(caption = "Select results file: "))
```
<br> 
<br>
Here we start the analysis:
* Unique_ID will identify a specific cell of 1 specific Condition 
* X_Shift and Y_Shift calculate the shift in X and Y, respectively, for each consecutive timepoint 
* Dist is the distance traveled per timepoint 
* Instant_Veloc is the istantaneous velocity (per each timepoint) 

```{r}
Data <- Data %>%
        #mutate(Unique_ID = paste0(Condition, ID)) %>%
        mutate(Time = Slice*Time_Res) %>%
        group_by(Track) %>%
        mutate(X_Shift = c(0,diff(X))^2,
               Y_Shift = c(0,diff(Y))^2,
               Dist = sqrt(X_Shift + Y_Shift),
               Instant_Veloc = Dist/unique(Time[2]),
               X_Shift = X - X[1],
               Y_Shift = Y - Y[1])

```
<br>
<br>
* Tot_Dist is the total migration distance
* Mean_Speed is the average migration speed
* Persistence is the "directionality" of the cell -> 1 = totally linear 0 = very unlinear
```{r}
Data_agg <- Data %>%
            group_by(Track)%>%
            summarise(Tot_Dist = sum(Dist, na.rm = T),
                      Mean_Speed = mean(Instant_Veloc, na.rm = T),
                      X_Displ = (first(X)-last(X))^2,
                      Y_Displ = (first(Y)-last(Y))^2,
                      Displ = sqrt(X_Displ + Y_Displ),
                      Persistence = (Displ/Tot_Dist))
```
<br>
<br>
Data_Avg_Speed will be used to compute the average speed, grouped per condition
x_limit and y_limit are used to define the limits for the full tracks plot

```{r}
Data_Avg_Speed <- Data %>%
  group_by(Time) %>% 
  summarise(Avg_Speed = mean(Instant_Veloc))

x_limit = ifelse(abs(min(Data$X_Shift)) > abs(max(Data$X_Shift)), abs(min(Data$X_Shift)), abs(max(Data$X_Shift)))
y_limit = ifelse(abs(min(Data$Y_Shift)) > abs(max(Data$Y_Shift)), abs(min(Data$Y_Shift)), abs(max(Data$Y_Shift)))
```
<br> 
<br>
Then we create all the plots
- Total distance and mean speed will show basically the same information but with different units
- Persistence will show the directionality of cell migration
- Instantaneous velocity shows how migration velocity changes over time
- Tracks_Lines_Centered shows the full track of all single cells, centered to the origin. This plot is useful to understand if the   migration is random or directed to a specific direction (persistence helps too). 


```{r}
Data_agg <- Data_agg %>%
            mutate(Condition = "Cells") %>% 
            group_by(Condition)%>%
            mutate(Avg_Speed = mean(Mean_Speed, na.rm = T),
                   sd_Speed = sd(Mean_Speed, na.rm = T),
                   Avg_Dist = mean(Tot_Dist, na.rm = T),
                   sd_Dist = sd(Tot_Dist, na.rm = T),
                   Avg_Pers = mean(Persistence, na.rm = T),
                   sd_Pers = sd(Persistence, na.rm = T),
                   )
```

```{r}

Tot_Dist_Boxplot <- ggplot(Data_agg, aes(x = Condition, y = Tot_Dist,color=as.factor(Track), fill = Condition)) +
                      #geom_boxplot(notch = F, alpha=0.7) +
                      geom_jitter(height = 0, width = 0.1, alpha=0.7, size = 3) +
                      geom_point(aes(x = Condition, y = Avg_Dist), color ="black", size = 5)+
                      geom_errorbar(aes(ymin = Avg_Dist-sd_Dist, ymax = Avg_Dist+sd_Dist), color ="black", width=.1) +
                      theme_bw()+
                      theme(axis.text.x = element_blank(),
                      axis.title.x = element_blank(),
                      axis.ticks.x = element_blank())+
                      ylab('Total Distance (um)')+
                      theme(axis.text.x = element_text(size = 16),
                      axis.text.y = element_text(size = 16),
                      axis.title.x = element_text(size = 20),
                      axis.title.y = element_text(size = 20),
                      legend.title = element_blank())


Mean_Speed_Boxplot <- ggplot(Data_agg, aes(x = Condition, y = Mean_Speed, fill = Condition,color=as.factor(Track))) +
                      #geom_boxplot(notch = F, alpha=0.7) +
                      geom_jitter(height = 0, width = 0.1, alpha=0.7, size = 3) +
                      geom_point(aes(x = Condition, y = avg), color ="black", size = 5)+
                      geom_errorbar(aes(ymin = avg-sd, ymax = avg+sd), color ="black", width=.1) +
                      theme_bw()+
                      theme(axis.text.x = element_blank(),
                      axis.title.x = element_blank(),
                      axis.ticks.x = element_blank())+
                      ylab('Mean Speed (um/min)')+
                      theme(axis.text.x = element_text(size = 16),
                      axis.text.y = element_text(size = 16),
                      axis.title.x = element_text(size = 20),
                      axis.title.y = element_text(size = 20),
                      legend.title = element_blank())
                    

Persistence_Boxplot <- ggplot(Data_agg, aes(x = Condition, y = Persistence,color=as.factor(Track),fill = Condition)) +
                        #geom_boxplot(notch = F, alpha=0.7) +
                        geom_jitter(height = 0, width = 0.1, alpha=0.7, size = 3) +
                        geom_point(aes(x = Condition, y = Avg_Pers), color ="black", size = 5)+
                        geom_errorbar(aes(ymin = Avg_Pers-sd_Pers, ymax = Avg_Pers+sd_Pers), color ="black", width=.1) +
                        theme_bw()+
                        theme(axis.text.x = element_blank(),
                        axis.title.x = element_blank(),
                        axis.ticks.x = element_blank())+
                        ylab('Directionality ratio')+
                        theme(axis.text.x = element_text(size = 16),
                        axis.text.y = element_text(size = 16),
                        axis.title.x = element_text(size = 20),
                        axis.title.y = element_text(size = 20),
                        legend.title = element_blank())
                   

Inst_Veloc_Line <- ggplot(Data, aes(x = Time, y = Instant_Veloc, color = as.factor(Track)))+ 
                    geom_line(aes(group = Track))+
                    geom_line(data=Data_Avg_Speed, aes(x = Time, y = Avg_Speed), color = 'red', size = 1.5)+
                    theme_bw()+
                    xlab('Time (min)')+
                    ylab('Instantaneous Velocity (um/min)')+
                    theme(axis.text.x = element_text(size = 16),
                    axis.text.y = element_text(size = 16),
                    axis.title.x = element_text(size = 20),
                    axis.title.y = element_text(size = 20),
                    legend.title = element_blank())
                    

Tracks_Path_Centered <- ggplot(Data, aes(x = X_Shift, y = Y_Shift, color=as.factor(Track)))+ 
                          geom_path(aes(group = Track), size = 1)+
                          geom_point(x=0,y=0, col = "red", size = 3)+
                          theme_bw()+
                          xlab('X (um)')+
                          ylab('Y (um)')+
                          theme(axis.text.x = element_text(size = 16),
                          axis.text.y = element_text(size = 16),
                          axis.title.x = element_text(size = 20),
                          axis.title.y = element_text(size = 20),
                          legend.title = element_blank())+
                          xlim(-400, 400)+
                          ylim(-400, 400)
```
<br>
<br>
Here we display all the plots
```{r}
Tot_Dist_Boxplot
#Mean_Speed_Boxplot
Persistence_Boxplot
Inst_Veloc_Line
Tracks_Path_Centered

```
<br>
<br>
And then, if we want (just uncomment the code), we save all the plots in a new folder in PDF format
```{r}
dir.create("../Plots")
setwd("../Plots")


Plots_List = list(Tot_Dist_Boxplot,
                  Mean_Speed_Boxplot,
                  Persistence_Boxplot,
                  Inst_Veloc_Line,
                  Tracks_Lines_Centered)

names(Plots_List) <- c("Tot_Dist_Jitter",
                       "Mean_Speed_Jitter",
                       "Persistence_Jitter",
                       "Inst_Veloc_Line",
                       "Tracks_Lines_Centered")
lapply(names(Plots_List),
       function(x)
         ggsave(filename=paste(x,".pdf",sep=""), height = 7, width = 7, plot=Plots_List[[x]]))
```

Here, we render and save an animated gif plot that reveals cell tracks over time, transitioning by time
```{r}
Tracks_Point_Centered <- ggplot(Data, aes(x = X_Shift, y = Y_Shift, color=as.factor(Track)))+ 
                          geom_path(aes(group = Track), size = 1)+
                          geom_point(x=0,y=0, col = "black", size = 5)+
                          theme_bw()+
                          xlab('X (um)')+
                          ylab('Y (um)')+
                          theme(axis.text.x = element_text(size = 16),
                          axis.text.y = element_text(size = 16),
                          axis.title.x = element_text(size = 20),
                          axis.title.y = element_text(size = 20),
                          legend.title = element_blank())+
                          xlim(-400, 400)+
                          ylim(-400, 400)

Tracks_Point_Centered + 
  transition_reveal(Time)
  #shadow_wake(wake_length = 0.5, alpha = FALSE)+
  #shadow_mark(size = 1)

anim_save(filename="Tracks_Point_Centered.gif",height = 7, width = 7)
```









